Incentivizing Exploration with Heterogeneous Utilities

### 2017-11-25 paper.tex
fixed a typo in algorithm 1 & edited the proof related to M

\begin{algorithm}
\caption{Algorithm: Incentivizing Exploration}
\label{Alg1}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\STATE Let $S = \{ i : P( \theta \cdot u_{i,t} > \theta \cdot u_{j,t}\ \forall j\ne i | u_{j,t}\ \forall j) < n^{-1}\}$ be the set of arms with unincentivized probability of being pulled below $n^{-1}$.
  \IF {$S$ is non-empty}
    \STATE{Choose an arm $i$ uniformly at random from $S$}
    \STATE{Pay whatever it takes to incentivize pulling arm $i$, i.e., offer payment 
    $c_{i,t} = \max_{\theta,j} \theta \cdot (u_{j,t} - u_{i,t})$ and $c_{j,t} = 0$ for $j \ne i$.}
  \ELSE
  \STATE {Let agents play myopically, i.e., offer payment $c_{j,t} = 0$ for all $j$}
  \ENDIF
  \STATE Update $u_{i,t}$ and $N(i,t)$ for the pulled arm
  \STATE Update the round number, $n = \min_{i} N(i,t)$
}\ENDFOR

@@

\begin{assumption} Let $f_{i,j}(y)$ be the marginal probability density function (or marginal probability mass function if $F(\cdot)$ is a discrete distribution) of $(u_i-u_j)\cdot\theta$ conditioned on $\theta \in \Omega_{i,j}$. We assume $M^{'}=\max_{i,j}\limsup_{y\rightarrow 0^{+}}f_{i,j}(y) <\infty$.
\label{A1}
\end{assumption}

Intuitively, assumption~\ref{A1} states that there are not many agents who are indifferent between their best arm and the second best arm.

@@

\begin{lemma}
For $t\in[t_n,t_{n+1}]$, the expected incentive payment $E[c(t)]$ is bounded above by $\Delta + 2WNm \sqrt{\frac{2}{n\pi}}$. 
\label{lemma:payment}
\end{lemma}

\begin{proof}
Denote $\Delta=\sup_{\theta}\max_{i}\{\theta\cdot u_{B(\theta)}-\theta\cdot u_{i}\}$ to be the theoretical maximum payment with no noisy. We consider the expected payment in the $n^{th}$ round. Denote $\epsilon_{i,t}=u_{i,t}-u_{i}$.

\begin{align}
&E[c(t)] \nonumber \\
\leq & \Delta \times P(c(t)>0) + E[\max_{i}\theta_t\cdot \epsilon_{i,t}-\min_{j}\theta_t\cdot \epsilon_{j,t}1\{c(t)>0\}] \nonumber \\
\leq & \Delta + E[\max_{i}\theta_t\cdot \epsilon_{i,t}-\min_{j}\theta_t\cdot \epsilon_{j,t}] \nonumber \\
\leq & \Delta + \sum_{j=1}^{m}2WE[\max_{i}|\epsilon_{i,t}^{j}|] \nonumber \\
\leq & \Delta + 2W\sum_{j=1}^{m}E\left[\sum_{i=1}^{N}|\epsilon_{i,t}^{j}|\right] \nonumber \\
\leq & \Delta + 2WNm E[|\epsilon_{1,t}^{1}|] \nonumber \\
\leq & \Delta + 2WNm \sqrt{\frac{2}{n\pi}} \nonumber
\end{align}

@@

Now we are ready to prove our first main result, Theorem~\ref{rst:budget}. \bccomment{The current proof about the cumulative expected payment is wrong. We need to bound the $E[c(t)|\textit{our algorithm decides to incentivize}]$ instead of $E[c(t)]$. However, the analysis for the cumulative regret is correct}

\begin{proof}
If $|u_{i}^{j}-u_{i,t}^{j}|\leq \lambda$ is true $\forall i$, $\forall j$, then we know for those $\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq 2m\lambda\}$, they will correctly identify their best arm. Thus we know, in the $n^{th}$ round, if $|u_{i}^{j}-u_{i,t}^{j}|\leq \frac{p^{-1}(\frac{p}{2})}{2m}$ $\forall i$ and $\forall j$, and $n^{-1}\leq p/2$, we do not need to incentivize any arms. In order to have $n^{-1}\leq \frac{p}{2}$, we need $n\geq \frac{2}{p}$. Denote $n_1=\max\{n_{0}, \frac{2}{p}\}$. Denote $\delta_{0}=p^{-1}(\frac{p}{2})>0$ (because of Assumption~\ref{A1}).The cumulative payment is bounded above by:
\begin{align}
&E[C(T)] \nonumber \\
= & \sum_{t=1}^{\infty}E[c(t)\mathbbm{1}\{\text{need to offer payment to an arm}\}] \nonumber \\
\leq &\sum_{n=1}^{\infty}\sum_{t=1}^{\infty}E[c(t)\mathbbm{1}\{t_{n-1}\leq t\leq t_{n}: \text{myopic action occurs regret for agents with } \nonumber \\
&\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq p^{-1}(n^{-1})\}\}] \nonumber \\
\leq &\sum_{n=n_{1}}^{\infty}\sum_{t=1}^{\infty}\left(\Delta+Nm\sqrt{\frac{2}{n\pi}}\right)P(\text{myopic action occurs regret for agents with }\nonumber \\
&\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq \delta_{0}\}|t_{n-1}\leq t\leq t_{n})P(t_{n-1}\leq t\leq t_{n})+\sum_{n=1}^{n_1}N\left(\Delta+\sqrt{\frac{2}{n\pi}}\right) \nonumber \\
\leq & \sum_{n=n_{1}}^{\infty}24Nm \exp\left(\frac{-1.8 n\delta_{0}^2}{32m^2\sigma^2}\right)\times Nn\left(\Delta+Nm\sqrt{\frac{2}{n_1\pi}}\right)+\sum_{n=1}^{n_1}N\left(\Delta+\sqrt{\frac{2}{n\pi}}\right) \nonumber 
\end{align}

We know the expected payment is bounded above by:
\begin{align}
E[C(T)] \leq 24N^2 m \frac{e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}}{(e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}-1)^2}\left(\Delta+Nm\sqrt{\frac{2}{n_1\pi}}\right)+\sum_{n=1}^{n_1}N\left(\Delta+\sqrt{\frac{2}{n\pi}}\right). \nonumber
\end{align}

Since we have $N$ arms in total, we know the $\mathbb{E}[C(T)]=O(N^3)$.
\end{proof}

Based on the proof of Theorem~\ref{rst:budget}, we know the expected number of payments we need to make is less than
\begin{align}
&\sum_{n=n_{1}}^{\infty}24Nm \exp\left(\frac{-1.8 n\delta_{0}^2}{32m^2\sigma^2}\right)\times Nn+Nn_{1} \nonumber \\
=&24N^2 m \frac{e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}}{(e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}-1)^2} + Nn_1, \nonumber
\end{align}
which we will use in the proof of Theorem~\ref{rst:regret}. Now we are ready to prove our second main result, Theorem~\ref{rst:regret}.

@@

Thus, the cumulative expected regret occured up to time $T$ when we let the agent play myopically is bounded above by:
\begin{align}
&E\left[\sum_{t=1}^{T}r(t)\right] \nonumber \\
=&E\left[\sum_{t=1}^{t_{n_{0}}} r(t) + \sum_{t=t_{n_{0}}}^{T}(r_1(t)+r_2(t))\right]  \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + E\left[\sum_{t=t_{n_{0}}}^{T}r_1(t)\right]+ E\left[\sum_{t=t_{n_{0}}}^{T}r_2(t)\right] \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + E\left[\sum_{n=n_{0}}^{T}\sum_{t=t_{n}}^{t_{n+1}}r_1(t)\right]+ E\left[\sum_{t=1}^{T}r_2(t)\right] \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=n_{0}}^{T}E\left[\sum_{t=t_{n}}^{t_{n+1}}24Nm\exp\left(-\frac{1.8n f(t)^2}{32 m^2\sigma^2}\right) R\right]+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=n_{0}}^{T}E\left[24Nm\exp\left(-\frac{1.8n f(t_n)^2}{32 m^2\sigma^2}\right)R \times (t_{n+1}-t_{n})\right]+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T}E\left[24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)R\times (t_{n+1}-t_{n})\right]+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \nonumber
\end{align}

The expected regret occurred when we incentivize the agents is calculated in Theorem~\ref{rst:budget}. Thus the cumulative regret at time $T$ is bounded above by
\begin{align}
&\sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
+ & 24N^2 m \frac{e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}}{(e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}-1)^2}+N\left(\max\left\{n_{0},\frac{2}{p}\right\}\right). \nonumber
\end{align}

@@

If we set $f^2(t)=\frac{2\log(T)\times 32m^2\sigma^2}{1.8n}$, then
\begin{align}
&\sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\ 
\leq & \sum_{n=1}^{T} 24N^2 mnR \exp\left(-2\log(T)\right)  + \frac{64m^2\sigma^2 M\log(T)}{1.8}\sum_{t=1}^{T}\frac{1}{n} \nonumber \\
\leq &  24N^2 m R\frac{T(T-1)}{2T^2}  + 35.56 m^2\sigma^2 M\log(T)(\log(T)+1) \nonumber \\
\leq &  12 N^2 m R  + 35.56 m^2\sigma^2 M\log(T)(\log(T)+1). \nonumber
\end{align}

@@

\subsection{Special Case}

In this section, we consider a special case of our problem, in which the agent's preference distribution $F(\cdot)$ satisfies $S(\delta)=0$ for $\delta<\delta^{'}$. The proof is similar to the proof of Theorem~\ref{rst:regret}, but for $t$ large enough, $S(f(t))=0$. In this case, with $O(N^3)$ payment budget, we can achieve $O(N^2)$ regret.

@@

\begin{algorithm}
\caption{Algorithm: Incentivizing Exploration Revised}
\label{Alg2}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\IF{the probability of pulling an arm $i$ is less than $n^{-1}$ and no agents have pulled that arm in this round (break ties randomly)}
	\STATE{Offer payment $c_{i,t}=\inf\{c: P(\theta\sim F: \theta\cdot u_{i,t}+c>max_{j}\theta\cdot u_{j,t})>n^{-1}\}$}\ELSE 
		\STATE {let the agents play myopically}\ENDIF
		\STATE $n = min_{i} N(i,t)$
}\ENDFOR

@@

In this section, we assume $\theta$ follows a continuous distribution $F(\cdot)$. We provide a example to show the best possible lower bound is $O(N\log(T))$.

>>>>

\begin{algorithm}
\caption{Algorithm: Incentivizing Exploration}
\label{Alg1}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number; Let $V=\emptyset$ be the set of arms that were pulled in the current round;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\STATE Let $S = \{ i : P( \theta \cdot u_{i,t} > \theta \cdot u_{j,t}\ \forall j\ne i | u_{j,t}\ \forall j) < n^{-1}\}$ be the set of arms with unincentivized probability of being pulled below $n^{-1}$.
  \IF {$S\setminus V$ is non-empty}
    \STATE{Choose an arm $i$ uniformly at random from $S\setminus V$}
    \STATE{Pay whatever it takes to incentivize pulling arm $i$, i.e., offer payment 
    $c_{i,t} = \max_{\theta,j} \theta \cdot (u_{j,t} - u_{i,t})$ and $c_{j,t} = 0$ for $j \ne i$.}
  \ELSE
  \STATE {Let agents play myopically, i.e., offer payment $c_{j,t} = 0$ for all $j$}
  \ENDIF
  \STATE Denote $A_t$ as the pulled arm, update $V=V\cup\{A_t\}$, $u_{A_t,t}$ and $N(A_t,t)$
  \IF {$n\neq \min_{i}N(i,t)$} 
  	\STATE $V=\emptyset$
  \ENDIF
  \STATE Update the round number, $n = \min_{i} N(i,t)$
}\ENDFOR

@@

\begin{assumption} Let $f_{i,j}(y)$ be the marginal probability density function (or marginal probability mass function if $F(\cdot)$ is a discrete distribution) of $(u_i-u_j)\cdot\theta$ conditioned on $\theta \in \Omega_{i,j}$. We assume $\max_{i,j}\sup_{y}f_{i,j}(y)\leq M$. 
\label{A1}
\end{assumption}

As we can see later in our proof, we only need $\max_{i,j}\limsup_{y\rightarrow 0^{+}}f_{i,j}(y)$ to be finite. Intuitively, assumption~\ref{A1} states that there are not many agents who are indifferent between their best arm and the second best arm.

@@

Now we are ready to prove our first main result, Theorem~\ref{rst:budget}. \bccomment{The proof below only bounds the cumulative expected number of payments}

\begin{proof}
If $|u_{i}^{j}-u_{i,t}^{j}|\leq \lambda$ is true $\forall i$, $\forall j$, then we know for those $\theta\in \{\theta:\theta\cdot u_{B(\theta)}-\max_{j\neq B(\theta)}\{\theta \cdot u_{j}\}\geq 2m\lambda\}$, they will correctly identify their best arm. Thus we know, in the $n^{th}$ round, if $|u_{i}^{j}-u_{i,t}^{j}|\leq \frac{p^{-1}(\frac{p}{2})}{2m}$ $\forall i$ and $\forall j$, and $n^{-1}\leq p/2$, we do not need to incentivize any arms. In order to have $n^{-1}\leq \frac{p}{2}$, we need $n\geq \frac{2}{p}$. Denote $n_1=\max\{n_{0}, \frac{2}{p}\}$. Denote $\delta_{0}=p^{-1}(\frac{p}{2})>0$ (because of Assumption~\ref{A1}).The cumulative expected number of payments is bounded above by:
\begin{align}
&E[\sum_{t=1}^{\infty}\mathbbm{1}\{c(t)>0\}] \nonumber \\
\leq &\sum_{n=n_{1}}^{\infty}\sum_{t=1}^{\infty}P\left(\exists i,j:|u_{i}^{j}-u_{i,t}^{j}|>\frac{p^{-1}(\frac{p}{2})}{2m}\right)+\sum_{n=1}^{n_1}N \nonumber \\
\leq & \sum_{n=n_{1}}^{\infty}24Nm \exp\left(\frac{-1.8 n\delta_{0}^2}{32m^2\sigma^2}\right)\times Nn+\sum_{n=1}^{n_1}N \nonumber  \\
=&24N^2 m \frac{e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}}{(e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}-1)^2} + Nn_1, \nonumber
\end{align}

Thus, we know the expected number of payments is bounded above by $O(N^2)$.

@@

Now we are ready to prove our second main result, Theorem~\ref{rst:regret}.

@@

Thus, the cumulative expected regret occured up to time $T$ when we let the agent play myopically is bounded above by:
\begin{align}
&E\left[\sum_{t=1}^{T}r(t)\right] \nonumber \\
=&E\left[\sum_{t=1}^{t_{n_{0}}} r(t) + \sum_{t=t_{n_{0}}}^{T}(r_1(t)+r_2(t))\right]  \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + E\left[\sum_{t=t_{n_{0}}}^{T}r_1(t)\right]+ E\left[\sum_{t=t_{n_{0}}}^{T}r_2(t)\right] \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + E\left[\sum_{n=n_{0}}^{T}\sum_{t=t_{n}}^{t_{n+1}}r_1(t)\right]+ E\left[\sum_{t=1}^{T}r_2(t)\right] \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=n_{0}}^{T}E\left[\sum_{t=t_{n}}^{t_{n+1}}24Nm\exp\left(-\frac{1.8n f(t)^2}{32 m^2\sigma^2}\right) R\right]+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=n_{0}}^{T}E\left[24Nm\exp\left(-\frac{1.8n f(t_n)^2}{32 m^2\sigma^2}\right)R \times (t_{n+1}-t_{n})\right]+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T}E\left[24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)R\times (t_{n+1}-t_{n})\right]+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
\leq & \sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \label{equ:regret}
\end{align}

The expected regret occurred when we incentivize the agents is calculated in Theorem~\ref{rst:budget}. Thus the cumulative regret at time $T$ is bounded above by
\begin{align}
&\sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\
+ & 24N^2 m \frac{e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}}{(e^{\frac{1.8\delta_{0}}{32m^2\sigma^2}}-1)^2}R+N\left(\max\left\{n_{0},\frac{2}{p}\right\}\right)R. \nonumber
\end{align}

@@

If we set $f^2(t)=\frac{2\log(T)\times 32m^2\sigma^2}{1.8t}$, then
\begin{align}
&\sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{T}Mf(t)^2 \nonumber \\ 
\leq & \sum_{n=1}^{T} 24N^2 mnR \exp\left(-2\log(T)\right)  + \frac{64m^2\sigma^2 M\log(T)}{1.8}\sum_{t=1}^{T}\frac{1}{n} \nonumber \\
\leq &  24N^2 m R\frac{T(T-1)}{2T^2}  + 35.56 m^2\sigma^2 M\log(T)(\log(T)+1) \nonumber \\
\leq &  12 N^2 m R  + 35.56 m^2\sigma^2 M\log(T)(\log(T)+1). \nonumber
\end{align}

@@

\begin{corollary}
If $\exists \delta>0$ such that $f_{i,j}(y)=0$ for all $i,j$ and $y\in [0,\delta]$, then the cumulative expected regret is bounded by $O(N^2)$.
\end{corollary}

\begin{proof}
The proof of this corollary is similar to the proof of Theorem~\ref{rst:regret}. If we set $f^2(t)=\frac{2\log(T)\times 32m^2\sigma^2}{1.8t}$, then there exists a $t_{0}$ such that for $t>t_{0}$, $S(f(t))=0$. Thus, similar to equation~\eqref{equ:regret}, we know the cumulative expected regret when we let the agents play myopically is bounded above by
\begin{align}
\sum_{n=1}^{n_{0}}NRn + \sum_{n=1}^{T} 24Nm\exp\left(-\frac{1.8n f(n)^2}{32 m^2\sigma^2}\right)\times R \times Nn+ \sum_{t=1}^{t_{0}}Mf(t)^2 \nonumber
\end{align}
Therefore, based on the same analysis of Theorem~\ref{rst:regret}, we know the cumulative regret is bounded by $O(N^2)$.
\end{proof}

@@

\begin{algorithm}
\caption{Algorithm: Incentivizing Exploration}
\label{Alg2}
\begin{algorithmic}
\STATE Set n = 1 to denote the round number; Let $V=\emptyset$ be the set of arms that were pulled in the current round;
\FOR{ $t = 1, 2, 3, \cdots$}{
	\STATE Let $S = \{ i : P( \theta \cdot u_{i,t} > \theta \cdot u_{j,t}\ \forall j\ne i | u_{j,t}\ \forall j) < n^{-1}\}$ be the set of arms with unincentivized probability of being pulled below $n^{-1}$.
  \IF {$S\setminus V$ is non-empty}
    \STATE{Choose an arm $i$ uniformly at random from $S\setminus V$}
	\STATE{Offer payment $c_{i,t}=\inf\{c: P(\theta\sim F: \theta\cdot u_{i,t}+c>max_{j}\theta\cdot u_{j,t})>n^{-1}\}$}
  \ELSE
  \STATE {Let agents play myopically, i.e., offer payment $c_{j,t} = 0$ for all $j$}
  \ENDIF
  \STATE Denote $A_t$ as the pulled arm, update $V=V\cup\{A_t\}$, $u_{A_t,t}$ and $N(A_t,t)$
  \IF {$n\neq \min_{i}N(i,t)$} 
  	\STATE $V=\emptyset$
  \ENDIF
  \STATE Update the round number, $n = \min_{i} N(i,t)$
}\ENDFOR

@@

In this section, we assume $\theta$ follows a continuous distribution $F(\cdot)$. We provide a example to show the best possible lower bound is $O(N\log(T))$ regardless of the incentivizing strategy.
